---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


---
title: "PA1_template.Rmd"
author: "Azam Yahya"
date: "October 13, 2015"
output: html_document
---

## Loading and preprocessing the data
```{r}
activity <- read.csv("Z:/Ferguson 13 march/other assignment/DS/Coursera/Courses/Reproducible research/Assignment/Assignment 1/Data/activity.csv")
```
Process/transform the data

```{r}
activity$date <- as.Date(activity$date,"%d-%m-%y")
activity <- na.omit(activity)
```
##What is mean total number of steps taken per day?

Calculate the total number of steps taken per day
```{r}
step_sum <- aggregate(steps ~ date, activity, sum)
print(step_sum)
```
Make a histogram of the total number of steps taken each day
```{r}
hist(step_sum$steps, col = "blue")
```
Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(step_sum$steps)
median(step_sum$steps)
```

* The Mean of the total numner of steps taken per day is 10766.19
* The Median of the total numner of steps taken per day is 10765

##What is the average daily activity pattern?
Make a time series plot

```{r}
library(ggplot2)
max <- aggregate(steps ~ interval, activity, mean)
g <- ggplot(max,aes(interval, steps))
g <- g + geom_line()
print(g)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max[which.max(max$steps), ]
```
* Interval 835 has the highest average number of steps i.e 206.17


##Imputing missing values
Calculate and report the total number of missing values in the dataset

```{r}
activity <- read.csv("Z:/Ferguson 13 march/other assignment/DS/Coursera/Courses/Reproducible research/Assignment/Assignment 1/Data/activity.csv")
sum(is.na(activity$steps))
```
* There are 2304 missing values

Devise a strategy for filling in all of the missing values in the dataset

* My strategy uses the mean for that 5-minute interval

```{r}
library(zoo, quietly = TRUE)
activity$steps1 <- na.aggregate(activity$steps, by = activity$interval, FUN = mean)
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity2 <- activity[c(4,2,3)]
```
Make a histogram of the total number of steps taken each day
```{r}
step_sum2 <- aggregate(steps1 ~ date, activity2, sum)
hist(step_sum2$steps1, col = "orange")
```

```{r}
mean(step_sum2$steps1)
median(step_sum2$steps1)
```
* The mean is same i.e. 10766.19 after imputing the missing values.
* There is a difference of 1.9 steps between the median after imputing the missing values

##Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r}
activity2$date <- as.Date(activity2$date,"%d-%m-%y")
activity2["week_type"] <- NA
activity2$week_type <- weekdays(activity2$date)
activity2$week_type <- ifelse(activity2$week_type =="Saturday" |activity2$week_type == "Sunday","weekend","weekday" )
activity2$week_type <- as.factor(activity2$week_type)
```
Make a panel plot containing a time series plot

```{r}
library(lattice)
week <- aggregate(steps1 ~ interval + week_type, activity2, mean)
xyplot(week$steps1 ~ week$interval | week$week_type, layout = c(1,2), type = "l")
```

* There are more fluctuation during the weekends than weekdays. However, the highest numners of steps results during the weekdays bwtween the interval of 800 to 900. 


## Are there differences in activity patterns between weekdays and weekends?
